Human Judgement Forecasting of COVID-19 in the UK

### 2023-04-10 manuscript/manuscript.tex

Accurate forecasts of infectious diseases can aid public health decision-making. In the past, two studies found ensembles of human judgement forecasts of COVID-19 to show predictive performance comparable to ensembles of computational models, at least when predicting case incidences. We present a follow-up to a study conducted in Germany and Poland and investigate a novel joint approach to combine human judgement and epidemiological modelling.

@@

From May 24th to August 16th 2021, we elicited weekly one to four-week-ahead forecasts of cases and deaths from COVID-19 in the UK from a crowd of human forecasters. Forecasts were elicited as part of a tournament, the "UK Crowd Forecasting Challenge", and a median ensemble of all forecasts was submitted to the European Forecast Hub. Participants could use two distinct interfaces: in one, forecasters submitted a predictive distribution directly, in the other forecasters instead submitted a forecast of the effective reproduction number $R_t$ based on an estimated baseline. This was then used to forecast cases and deaths using simulation methods from the \texttt{EpiNow2} \textsf{R} package. Forecasts were scored using the weighted interval score on the original forecasts, as well as after applying the natural logarithm to both forecasts and observations.

@@

The ensemble of human forecasters overall performed comparably to the official European Forecast Hub ensemble on both cases and deaths, although results were susceptible to changes in details of the evaluation. $R_t$ forecasts performed comparably to direct forecasts on cases, but worse on deaths. Self-identified experts did not outperform non-experts though the sample size was small.

@@

Human judgement can be useful in predicting the spread of an infectious disease such as COVID-19. The results of forecast evaluations can change depending on what metrics are chosen and judgement on what does or doesn't constitute a "good" forecast is dependent on the forecast consumer. Combinations of human and computational forecasts hold potential but present real-world challenges that need to be solved.

@@

Past research found that in the context of infectious disease forecasting, human judgement forecasts could achieve predictive performance broadly comparable to forecasts generated based on mathematical modelling, in particular when forecasting incident cases. \citet{farrowHumanJudgmentApproach2017} found that an aggregate of human predictions outperformed computational models when predicting the 2014/15 and 2015/16 flu season in the US. However, a comparable approach performed worse than computational models at predicting the 2014/15 outbreak of chikungunya in the Americas. 
\citet{bosseComparingHumanModelbased2022} found an ensemble of human forecasters to outperform an ensemble of computational models when predicting cases of COVID-19 in Germany, but performing worse when predicting incident deaths. Similarly, \citet{mcandrewChimericForecastingCombining2022} reported an ensemble of human forecasters to perform comparably to an ensemble of computational models when predicting incident cases, and worse when predicting incident deaths. \citet{farrowHumanJudgmentApproach2017} and in particular \citet{bosseComparingHumanModelbased2022} struggled to recruit a large number of participants (numbers of active forecasters ranged from 22 to 61 in \citet{mcandrewChimericForecastingCombining2022}, 7 to 24 in \citet{farrowHumanJudgmentApproach2017}, and 4 to 10 in \citet{bosseComparingHumanModelbased2022}). 
It is important to note that in previous studies (and also this one) human forecasters were free to use any resources, including computational models, in the process of creating a forecast, making it difficult to completely separate human judgement and computational modelling.

In some situations, human judgement forecasting may have advantages relative to computational models. Human judgment may be particularly useful to provide timely forecasts in situations where data is sparse and many parameters are unknown. Humans are generally able to answer a broad set of question (such as for example the likelihood that a given actor will take some specified action) or can take factors into account that are hard to encode in a computational model. On the other hand, human judgement forecasting is difficult to scale due to the time and effort required, and humans may be at a disadvantage at tasks that strongly benefit from the ability to perform complex computations. Also, the use of human judgement forecasts by decision makers may be complicated by the lack of clarity of the basis on which they were made.

@@

Methods that aim to combine human judgement and mathematical modelling are therefore appealing, though we note that presenting this as a binary choice is misleading as most computational models in use in epidemiology have at least some element of human judgement supporting their structure or usage and many human forecasters make use of simple base rate based approaches which are in reality simple models. One explicit method to combine separate human judgement and computational model forecasts with the goal of improving predictive performance is an ensemble. This has been shown to improve performance across model types \citep{mcandrewChimericForecastingCombining2022}. \citet{farrowHumanJudgmentApproach2017, bosseComparingHumanModelbased2022, swallowChallengesEstimationUncertainty2022} and others suggested additional possibilities in the context of infectious diseases that may also help reduce the amount of human effort required. One approach is to use human forecasts, for example of relevant disease parameters, as an input to computational modelling. Another approach is to use mathematical modelling in explicit combination with human judgement, for example by giving experts the option to make post-hoc adjustments to model outputs. \citet{bosseComparingHumanModelbased2022} proposed asking human forecasters to forecast the effective reproduction number $R_t$ (the average number of people an infected person would infect in turn) based on modelled estimates and to then use this forecast in a mathematical simulation model in order to obtain forecasts for observed case and death numbers.

@@

The European COVID-19 Forecast Hub \cite{sherrattPredictivePerformanceMultimodel2022a} started in March 2021 to elicit predictions for various COVID-19 related forecast targets from different research groups every week. The forecasts evaluated in this study were submitted every Monday before 11.59pm GMT between May 24 2021 and August 16 2021. Forecasts were made for incident weekly reported numbers of cases of and deaths from COVID-19 on a national level for various European countries over a one to four week forecast horizon. While forecasts were submitted on Mondays, weeks were defined as epidemiological weeks, ending on a Saturday and starting on Sunday. Forecast horizons were therefore in fact 5, 12, 19 and 26 days. Submissions to the European Forecast Hub followed a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15,…, 0.95, 0.975, 0.99.
Every week, forecasts submitted to the hub were automatically checked for conformity with the required format and all eligible forecasts combined into different ensembles. Until the 12th of July 2021 the default Hub ensemble shown on all official Forecast Hub visualisations (https://covid19forecasthub.eu/) was a mean-ensemble (i.e., the $\alpha$-quantile of the ensemble is given by the mean of all submitted $\alpha$-quantiles). From the 29th of July onwards, the default Forecast Hub ensemble became a median ensemble. The median number of models included in the Forecast Hub ensemble during the study period was 9 for cases and 10 for deaths (see Figure \ref{fig:num-forecasters} in the SI).

Ground-truth data on daily reported test positive cases and deaths linked to COVID-19 were provided by the European Forecast Hub and sourced from the Johns Hopkins University (JHU). Data were subject to reporting artifacts and revisions. All data points were marked as anomalous retrospectively by the European Forecast Hub if in subsequent updates data was changed by more than 5 percent. In August 2022 JHU switched the data source for their death numbers from "deaths within 28 days of a positive COVID test" to "Deaths with COVID-19 on the death certificate" and revised all their past data to guarantee consistency. The 2021 UK ground truth death data as it was made available through the European Forecast Hub in 2021 is therefore substantially different and on average lower than the data available as of early 2023. The data and revisions are displayed in Figure \ref{fig:plot-data}. All results presented here were derived based on the original data available in 2021, which were available through the European COVID-19 Forecast Hub GitHub repository (\url{https://github.com/covid19-forecast-hub-europe/covid19-forecast-hub-europe}).

@@

When signing up, we asked participants to self-identify as "experts", asking them to tick a box if they worked in infectious disease modelling or had professional experience in any related field.

@@

Just as in the previous study, the default forecast shown was a repetition of the last known observation with constant uncertainty around it computed as the standard deviation of the last four changes in weekly log observed forecasts (i.e. as $\sigma(\log(\text{value4}) - \log(\text{value3}), \, \log(\text{value3}) - \log(\text{value2}), \, \dots)$).
In addition to information about past observations, participants could see various metrics and data such as the test positivity rate and vaccination rate sourced from Our World in Data \citep{owidcoronavirus}. Figure \ref{fig:screenshot-classical} shows a screenshot of the forecast interface for direct forecasts.

In addition to the 'direct' forecasts, we implemented a second forecasting method ('$R_t$ forecasts'), where we asked participants to make a forecast of the effective reproduction number $R_t$ based on a baseline estimate produced by the \texttt{EpiNow2} \citep{epinow2} \textsf{R} \cite{R} package effective reproduction number model which we also used in \cite{bosseComparingHumanModelbased2022} as a standalone computational model. The estimate produced by \texttt{EpiNow2} was shown as the default forecast and could be adjusted by the user. The resulting $R_t$ forecast was then translated into a forecast of cases using the simulation model from the \texttt{EpiNow2} \textsf{R} package, which implements a renewal equation based \citep{fraserEstimatingIndividualHousehold2007} generative process for latent infections. We chose a Gaussian Process prior with mean 0 for the first differences of the effective reproduction number in time, implying that in the absence of informative data the reproduction number would remain constant on average, with uncertainty increasing with the temporal distance to informative data points. Latent infections were convolved with delay distributions representing the incubation period and reporting delay, and assumed to follow a negative binomial observation model with a day of the week effect to produce an estimate of reported cases. This approach has been widely used for short-term forecasting \cite{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a} and used to produce reproduction number estimates \cite{abbottEstimatingIncreaseReproduction2021, abbottEstimatingTimevaryingReproduction2020a, sherrattExploringSurveillanceData2021}. Further details are given in the SI.

@@

Every week, we submitted an ensemble of individual forecasts to the European Forecast Hub. In contrast to the ensemble of human forecasts described in \citet{bosseComparingHumanModelbased2022}, we used the quantile-wise median, rather than the quantile-wise mean to combine predictions. %, following suggestions made in CITATION. 
We submitted three different ensembles to the Hub: The first one, "epiforecasts-EpiExpert\_direct" (here called "direct crowd forecast" or "crowd-direct") was a quantile-wise median ensemble of all the direct forecasts. "epiforecasts-EpiExpert\_Rt" (here called "$R_t$ forecast" or "crowd-rt") was a median ensemble of all forecasts made through the $R_t$ interface. "epiforecasts-EpiExpert" (here called "combined crowd ensemble" or "crowd-ensemble") was a median ensemble of all forecasts together. A participant could enter the combined crowd ensemble twice if they had submitted both a direct and an $R_t$ forecast. Before creating the ensemble, we deleted forecasts that were clearly the result of a user or software error (such as forecasts that were zero everywhere).

@@

In terms of WIS on the log scale, all human forecasting approaches outperformed the Forecast Hub ensemble for two week ahead forecasts of cases (see Figure \ref{fig:performance}). WIS values relative to the Hub ensemble (=1) were 0.91 for the combined crowd ensemble, 0.96 for the direct crowd forecasts and 0.93 for the $R_t$ forecasts (see Table \ref{tab:scores}). 
In terms of WIS on the natural scale, however, the Hub ensemble was ahead of all human forecasting approaches. Relative WIS values on the natural scale for two week ahead forecasts were 1.05 for the combined crowd ensemble, 1.03 for the direct crowd forecasts and 1.2 for the $R_t$ forecasts.

@@

In terms of relative model ranks for two week ahead forecasts, the Hub ensemble and the $R_t$ showed a higher variance than the combined crowd ensemble and the direct forecasts (See Figure \ref{fig:performance-ranks}). Both the Hub ensemble and the $R_t$ forecast were more often in first place than other approaches (4 times each, both on the log and on the natural scale). However, they were also most often in the last place (Hub ensemble: 6 on the log scale and 5 on the natural scale, $R_t$: 5 on the log scale and 6 on the natural scale). The direct forecasts placed relatively equally in places 1-4. The crowd ensemble never placed fourth, but also had the lowest number of first places (2, both on the log and the natural scale). Model ranks only changed marginally when switching between the log and the natural scale.

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/performance-ranks.png}
\caption{\bf{Performance ranks}....}
\label{fig:performance-ranks}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/performance-ranks-4.png}
\caption{\bf{Performance ranks}....}
\label{fig:performance-ranks-4}
\end{figure*}

@@

All forecasting approaches except the $R_t$ forecasts showed higher empirical coverage for deaths than for cases (see Figure \ref{fig:performance}). For 50\% prediction intervals, the Hub ensemble exceeded the nominal coverage noticeably (0.77) (see Table \ref{tab:scores}). $R_t$ forecasts failed to get close to nominal coverage (0.15), while the combined crowd ensemble and the direct forecasts had empirical coverage close to nominal coverage (both 0.54). For 90\% prediction intervals, the Hub ensemble again exceeded nominal coverage and covered all observations (1) while the $R_t$ forecasts again failed to get close to nominal coverage (0.46). The crowd ensemble exhibited some underdispersion (0.77) while the direct forcecasts almost reached nominal coverage for two week ahead forecasts of deaths (0.85).

In terms of WIS on the log scale for two week ahead predictions of deaths, the combined crowd ensemble (0.07) and the direct crowd forecasts (0.99) were slightly ahead of the Hub ensemble, while the $R_t$ forecasts performed noticeably worse (1.98) (see Figure \ref{fig:performance} and Table \ref{tab:scores}). Interestingly, combining the $R_t$ forecasts and the direct forecasts led to an ensemble that performed better than either of them. 
In terms of WIS on the natural scale, only the direct forecasts (0.89) performed better for two week ahead death predictions than the Hub ensemble, while the combined crowd ensemble performed a bit worse (1.06) and the $R_t$ forecasts again noticeably worse (2.1).

@@

When comparing scores on the log and on the natural scale, scores on the log scale were again more evenly distributed across the study period. On the natural scale, high scores were concentrated around the end of the study period, were death incidences were highest (see Figure \ref{fig:forecasts-scores}).

\subsection*{Rt forecasts}

@@

%Mention calbration

@@

A median ensemble of two week ahead forecasts made only by "experts" or "non-experts" (determined based on self-reported experience in infectious disease modelling or a related field) performed worse than the combined crowd example, both for cases and deaths and both on the log scale and on the natural scale (see Figure \ref{fig:performance-experts} and Table \ref{tab:scores-experts}).

When comparing "experts" and "non-experts", "experts" were better calibrated for forecasts two weeks ahead (see Figure \ref{fig:performance-experts}. For cases two week ahead, "experts" achieved better scores than "non-experts" both on the log and on the natural scale. WIS values \textit{relative to the combined crowd ensemble} were 1.08 for "experts" and 1.14 for "non-experts" on the log scale and 1.06 for "experts" and 1.07 for "non-experts" on the natural scale (see Table \ref{tab:scores-experts}).

For two week ahead forecasts of deaths, "experts" performed worse than "non-experts" in terms of WIS on the log scale (WIS relative to the combined crowd ensemble: 1.29 vs. 1.06), but better on the natural scale (1.03 vs. 1.14).

@@

% \paragraph{Effort to recruit participants}. 
Just like \citet{bosseComparingHumanModelbased2022} and \citet{farrowHumanJudgmentApproach2017}, this study struggled to retain a large number of participants. Focused public outreach efforts such as creating a dedicated website, announcing an official tournament, providing a public leaderboard, sending weekly emails with details on past performance and weekly announcements on Twitter, did noticeably increase participation compared to the previous study in Germany and Poland. Nevertheless, retaining participants beyond the initial recruitment proved challenging, and most forecasters only submitted a single forecast. \citet{mcandrewChimericForecastingCombining2022} had a higher number of participants, suggesting that making use of existing forecasting platforms, such as Metaculus or Good Judgement Open that access to a large existing user base and greater resources may be helpful in recruiting a larger number of participants though these lack the flexibility and software tooling to run a novel study of this kind in real-time as things stand.

@@

%WIS scores
For forecasts of cases two weeks ahead, performance of the human judgement forecasts was better than the European Forecast Hub ensemble in terms of WIS on the log scale, and worse in terms of WIS on the natural scale. For forecasts of deaths two weeks ahead, direct human forecasts and the combined crowd ensemble performed better than the Hub ensemble on the log scale. On the natural scale, the combined crowd ensemble performed worse than the Hub ensemble, while the direct crowd forecasts still performed better. $R_t$ forecasts for deaths performed noticeably worse than all other approaches both on the log and on the natural scale.

In their original study conducted in Germany and Poland, \citet{bosseComparingHumanModelbased2022} humans outperformed an ensemble of computational models when predicting cases, but not when predicting deaths. They hypothesised that computational models might have an advantage over human forecasters when predicting deaths, benefiting from the ability to model the delays and epidemiological relationships between different leading and lagged indicators. \citet{mcandrewChimericForecastingCombining2022} similarly found in their study that humans performed comparably to an ensemble of computational models for cases, but not for predictions of deaths of COVID-19. 
Results in our study do not directly support this pattern, but given the low number of observations also do not provide strong evidence against it. In this study, the combined crowd ensemble performed better than the Hub ensemble on both cases and deaths on the log scale, on worse on the natural scale. Direct forecasts, which would be most comparable to the forecasts in \citep{bosseComparingHumanModelbased2022}, performed worse than the Hub ensemble on cases and better on deaths. 
Relatively good performance of direct human judgement forecasts for death forecasts compared to previous studies may be related to changes in the age distribution of cases and the rise of the Delta variant in the UK starting in the beginning of May. The Delta variant had a higher case fatality ratio (CFR) than previous variants \cite{linDiseaseSeverityClinical2021}. One possible hypothesis for the relatively good performance of human forecasts for deaths compared to previous studies might be that some models submitted to the Forecast Hub may have been more negatively affected by the changes in CFR during the study period than human forecasters.

>>>>

In the past, two studies found ensembles of human judgement forecasts of COVID-19 to show predictive performance comparable to ensembles of computational models, at least when predicting case incidences. We present a follow-up to a study conducted in Germany and Poland and investigate a novel joint approach to combine human judgement and epidemiological modelling.

@@

From May 24th to August 16th 2021, we elicited weekly one to four-week-ahead forecasts of cases and deaths from COVID-19 in the UK from a crowd of human forecasters. Forecasts were elicited as part of a tournament, the "UK Crowd Forecasting Challenge", and a median ensemble of all forecasts was submitted to the European Forecast Hub. Participants could use two distinct interfaces: in one, forecasters submitted a predictive distribution directly, in the other forecasters instead submitted a forecast of the effective reproduction number $R_t$. This was then used to forecast cases and deaths using simulation methods from the \texttt{EpiNow2} \textsf{R} package. Forecasts were scored using the weighted interval score on the original forecasts, as well as after applying the natural logarithm to both forecasts and observations.

@@

The ensemble of human forecasters overall performed comparably to the official European Forecast Hub ensemble on both cases and deaths, although results were sensitive to changes in details of the evaluation. $R_t$ forecasts performed comparably to direct forecasts on cases, but worse on deaths. Self-identified "experts" outperformed "non-experts" in terms of WIS on the log scale, but not on the natural scale.

@@

Human judgement forecasts and computational models can produce forecasts of similar quality for infectious disease such as COVID-19. The results of forecast evaluations can change depending on what metrics are chosen and judgement on what does or doesn't constitute a "good" forecast is dependent on the forecast consumer. Combinations of human and computational forecasts hold potential but present real-world challenges that need to be solved.

@@

Past research found that in the context of infectious disease forecasting, human judgement forecasts could achieve predictive performance broadly comparable to forecasts generated based on mathematical modelling, in particular when forecasting incident cases, rather than lagged indicators indicators like deaths. \citet{farrowHumanJudgmentApproach2017} found that an aggregate of human predictions outperformed computational models when predicting the 2014/15 and 2015/16 flu season in the US. However, a comparable approach performed worse than computational models at predicting the 2014/15 outbreak of chikungunya in the Americas. 
\citet{bosseComparingHumanModelbased2022} found an ensemble of human forecasters to outperform an ensemble of computational models when predicting cases of COVID-19 in Germany and Poland, but performing worse when predicting incident deaths. Similarly, \citet{mcandrewChimericForecastingCombining2022} reported an ensemble of human forecasters to perform comparably to an ensemble of computational models when predicting incident COVID-19 cases, and worse when predicting incident deaths. \citet{farrowHumanJudgmentApproach2017} and in particular \citet{bosseComparingHumanModelbased2022} struggled to recruit a large number of participants (numbers of active forecasters ranged from 22 to 61 in \citet{mcandrewChimericForecastingCombining2022}, 7 to 24 in \citet{farrowHumanJudgmentApproach2017}, and 4 to 10 in \citet{bosseComparingHumanModelbased2022}). 
It is important to note that in previous studies (and also this one) human forecasters were free to use any resources, including computational models, in the process of creating a forecast, making it difficult to completely separate human judgement and computational modelling.

In some situations, human judgement forecasting may have advantages relative to computational models. Human judgment may be particularly useful to provide timely forecasts in situations where data is sparse and many parameters are unknown or structural breaks have occurred. Humans are generally able to answer a broad set of question (such as for example the likelihood that a given actor will take some specified action) or can take factors into account that are hard to encode in a computational model. On the other hand, human judgement forecasting is difficult to scale due to the time and effort required, and humans may be at a disadvantage at tasks that strongly benefit from the ability to perform complex computations. Also, the use of human judgement forecasts by decision makers may be complicated by the lack of clarity of the basis on which they were made.

@@

Methods that aim to combine human judgement and mathematical modelling are therefore appealing, though we note that presenting this as a binary choice is misleading. Most computational models in use in epidemiology have at least some element of human judgement supporting their structure or usage. Also, human forecasters often make use of approaches such as calculating a base rate of incidences, or extrapolating current trends, which are in reality equivalent to simple models. One explicit method to combine separate human judgement and computational model forecasts with the goal of improving predictive performance is an ensemble. This has been shown to improve performance across model types \citep{mcandrewChimericForecastingCombining2022}. \citet{farrowHumanJudgmentApproach2017, bosseComparingHumanModelbased2022, swallowChallengesEstimationUncertainty2022} and others suggested additional possibilities in the context of infectious diseases that may also help reduce the amount of human effort required. One approach is to use human forecasts, for example of relevant disease parameters, as an input to computational modelling. Another approach is to use mathematical modelling in explicit combination with human judgement, for example by giving experts the option to make post-hoc adjustments to model outputs. \citet{bosseComparingHumanModelbased2022} proposed asking human forecasters to forecast the effective reproduction number $R_t$ (the average number of people an infected person would infect in turn) based on modelled estimates and to then use this forecast in a mathematical simulation model in order to obtain forecasts for observed case and death numbers.

@@

The European COVID-19 Forecast Hub \cite{sherrattPredictivePerformanceMultimodel2022a} was launched in March 2021 in order to elicit weekly predictions for various COVID-19 related forecast targets from different research groups. The forecasts evaluated in this study were submitted every Monday before 11.59pm GMT between May 24 2021 and August 16 2021. Forecasts were made for incident weekly reported numbers of cases of and deaths from COVID-19 on a national level for various European countries over a one to four week forecast horizon. While forecasts were submitted on Mondays, weeks were defined as epidemiological weeks, ending on a Saturday and starting on Sunday. Forecast horizons were therefore in fact 5, 12, 19 and 26 days. Submissions to the European Forecast Hub followed a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15,…, 0.95, 0.975, 0.99.
Every week, forecasts submitted to the hub were automatically checked for conformity with the required format and all eligible forecasts combined into different ensembles. Until the 12th of July 2021 the default Hub ensemble shown on all official Forecast Hub visualisations (https://covid19forecasthub.eu/) was a mean-ensemble (i.e., the $\alpha$-quantile of the ensemble is given by the mean of all submitted $\alpha$-quantiles). From the 29th of July onwards, the default Forecast Hub ensemble became a median ensemble. The median number of models included in the Forecast Hub ensemble for the UK during the study period was 9 for cases and 10 for deaths (see Figure \ref{fig:num-forecasters} in the SI).

Ground-truth data on daily reported test positive cases and deaths linked to COVID-19 were provided by the European Forecast Hub and sourced from the Johns Hopkins University (JHU). Data were subject to reporting artifacts and revisions. All data points were marked as anomalous retrospectively by the European Forecast Hub if in subsequent updates data was changed by more than 5 percent. In August 2022 JHU switched the data source for their UK death numbers from "deaths within 28 days of a positive COVID test" to "Deaths with COVID-19 on the death certificate" and revised all their past data to guarantee consistency. The 2021 UK ground truth death data as it was made available through the European Forecast Hub in 2021 is therefore substantially different and on average lower than the data available as of early 2023. The data and revisions are displayed in Figure \ref{fig:plot-data}. All results presented here were derived based on the original data available in 2021, which were available through the European COVID-19 Forecast Hub GitHub repository (\url{https://github.com/covid19-forecast-hub-europe/covid19-forecast-hub-europe}).

@@

When signing up, participants could self-identify as "experts" if they worked in infectious disease modelling or had professional experience in any related field.

@@

Just as in the previous study, the default forecast shown was a repetition of the last known observation with constant uncertainty around it. The shown distribution was the exponential of a normal distribution with mean $\log (\text{last value})$ and uncertainty set to the standard deviation of the last four changes in weekly log observed forecasts (i.e. as $\sigma(\log(\text{value4}) - \log(\text{value3}), \, \log(\text{value3}) - \log(\text{value2}), \, \dots)$).
In addition to information about past observations, participants could see various metrics and data such as the test positivity rate and vaccination rate sourced from Our World in Data \citep{owidcoronavirus}. Figure \ref{fig:screenshot-classical} shows a screenshot of the forecast interface for direct forecasts.

In addition to the 'direct' forecasts, we implemented a second forecasting method ('$R_t$ forecasts'), where we asked participants to make a forecast of the effective reproduction number $R_t$. This forecast was made based on a baseline estimate produced by the \texttt{EpiNow2} \citep{epinow2} \textsf{R} \cite{R} package effective reproduction number model which we also used in \cite{bosseComparingHumanModelbased2022} as a standalone computational model. The estimate produced by \texttt{EpiNow2} was shown as the default forecast and could be adjusted by the user. The resulting $R_t$ forecast was then translated into a forecast of cases using the simulation model from the \texttt{EpiNow2} \textsf{R} package, which implements a renewal equation based \citep{fraserEstimatingIndividualHousehold2007} generative process for latent infections. We chose a Gaussian Process prior with mean 0 for the first differences of the effective reproduction number in time, implying that in the absence of informative data the reproduction number would remain constant on average, with uncertainty increasing with the temporal distance to informative data points. Latent infections were convolved with delay distributions representing the incubation period and reporting delay, and assumed to follow a negative binomial observation model with a day of the week effect to produce an estimate of reported cases. This approach has been widely used for short-term forecasting \cite{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a} and used to produce reproduction number estimates \cite{abbottEstimatingIncreaseReproduction2021, abbottEstimatingTimevaryingReproduction2020a, sherrattExploringSurveillanceData2021}. Further details are given in the SI.

@@

Every week, we submitted an ensemble of individual forecasts to the European Forecast Hub. In contrast to the ensemble of human forecasts described in \citet{bosseComparingHumanModelbased2022}, we used the quantile-wise median, rather than the quantile-wise mean to combine predictions, drawing upon insights gained from the COVID-19 Forecast Hubs \citep{rayComparingTrainedUntrained2022}. 
We submitted three different ensembles to the Hub: The first one, "epiforecasts-EpiExpert\_direct" (here called "direct crowd forecast" or "crowd-direct") was a quantile-wise median ensemble of all the direct forecasts. "epiforecasts-EpiExpert\_Rt" (here called "$R_t$ forecast" or "crowd-rt") was a median ensemble of all forecasts made through the $R_t$ interface. "epiforecasts-EpiExpert" (here called "combined crowd ensemble" or "crowd-ensemble") was a median ensemble of all forecasts together. A participant could enter the combined crowd ensemble twice if they had submitted both a direct and an $R_t$ forecast. Before creating the ensemble, we deleted forecasts that were clearly the result of a user or software error (such as forecasts that were zero everywhere).

@@

In terms of WIS on the log scale, all human forecasting approaches outperformed the Forecast Hub ensemble for two week ahead forecasts of cases (see Figure \ref{fig:performance}). WIS values relative to the Hub ensemble (=1) were 0.91 for the combined crowd ensemble, 0.96 for the direct crowd forecasts and 0.93 for the $R_t$ forecasts (see Table \ref{tab:scores}). 
In terms of WIS on the natural scale, however, the Hub ensemble was ahead of all human forecasting approaches. Relative WIS values on the natural scale for two week ahead forecasts were 1.05 for the combined crowd ensemble, 1.03 for the direct crowd forecasts and 1.2 for the $R_t$ forecasts. 
This discrepancy between performance on the log scale can be attributed to the fact that case forecasts of the Hub ensemble tended to be lower than forecasts of the human judgement forecasting approaches (see Figure \ref{fig:forecasts-scores}). On the natural scale, this led to smaller overprediction penalties on the natural scale, putting it ahead of human forecasts (see Figure \ref{fig:performance}A,C). On the log scale, however, it led to large penalties for underprediction.

@@

In terms of relative model ranks for two week ahead forecasts, the Hub ensemble and the $R_t$ showed a higher variance than the combined crowd ensemble and the direct forecasts (See Figure \ref{fig:performance-ranks}), despite forecasts being about the same or more dispersed (see Figure \ref{fig:performance}). Both the Hub ensemble and the $R_t$ forecast were more often in first place than other approaches (4 times each, both on the log and on the natural scale). However, they were also most often in the last place (Hub ensemble: 6 on the log scale and 5 on the natural scale, $R_t$: 5 on the log scale and 6 on the natural scale). The direct forecasts placed relatively equally in places 1-4. The crowd ensemble never placed fourth, but also had the lowest number of first places (2, both on the log and the natural scale). Model ranks only changed marginally when switching between the log and the natural scale.

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/performance-ranks.png}
\caption{\bf{Ranks for all forecasting approaches for two week ahead forecasts}. Colours indicate how often (out of 13 forecasts) a given approach got 1st, 2nd, 3rd, or 4th rank.}
\label{fig:performance-ranks}
\end{figure*}

@@

All forecasting approaches except the $R_t$ forecasts showed higher empirical coverage for deaths than for cases (see Figure \ref{fig:performance}). Forecasts from the Hub ensemble generally tended to be wider than the human forecasts (see Figures \ref{fig:forecasts-scores} and \ref{fig:performance}B,D). For 50\% prediction intervals, the Hub ensemble exceeded the nominal coverage noticeably (0.77) (see Table \ref{tab:scores}). $R_t$ forecasts failed to get close to nominal coverage (0.15), while the combined crowd ensemble and the direct forecasts had empirical coverage close to nominal coverage (both 0.54). For 90\% prediction intervals, the Hub ensemble again exceeded nominal coverage and covered all observations (1) while the $R_t$ forecasts again failed to get close to nominal coverage (0.46). The crowd ensemble exhibited some underdispersion (0.77) while the direct forcecasts almost reached nominal coverage for two week ahead forecasts of deaths (0.85).

In terms of WIS on the log scale for two week ahead predictions of deaths, the combined crowd ensemble (0.97) and the direct crowd forecasts (0.99) were marginally ahead of the Hub ensemble, while the $R_t$ forecasts performed noticeably worse (1.98) (see Figure \ref{fig:performance}D and Table \ref{tab:scores}). For the Hub ensemble, the dispersion component played by far the largest role, while this was less the case for the human forecasts, which got higher penalties from both over- and underprediction. Interestingly, combining the $R_t$ forecasts and the direct forecasts led to an ensemble that performed better than either of them alone on the log scale. 
In terms of WIS on the natural scale, only the direct forecasts (0.89) performed better for two week ahead death predictions than the Hub ensemble, while the combined crowd ensemble performed a bit worse (1.06) and the $R_t$ forecasts again noticeably worse (2.1).

@@

When comparing scores on the log and on the natural scale, scores on the log scale were again more evenly distributed across the study period. On the natural scale, high scores were concentrated around the end of the study period, when death incidences were highest (see Figure \ref{fig:forecasts-scores}).

\subsection*{$R_t$ forecasts}

@@

@@

A median ensemble of two week ahead forecasts restricted to only those made by either "experts" or "non-experts" (determined based on self-reported experience in infectious disease modelling or a related field) performed worse than the combined crowd example, both for cases and deaths and both on the log scale and on the natural scale (see Figure \ref{fig:performance-experts} and Table \ref{tab:scores-experts} and Figure \ref{fig:num-forecasters}B for a visualisation of participation). The median number of "non-experts" was 11 for cases and 10 for deaths, which was higher than the median number of "experts", which was 8 for cases and deaths.

When comparing "experts" and "non-experts", the ensemble of "experts" was better calibrated for forecasts two weeks ahead (see Figure \ref{fig:performance-experts}. For cases two week ahead, "experts" achieved better scores than "non-experts" both on the log and on the natural scale. WIS values \textit{relative to the combined crowd ensemble} were 1.08 for "experts" and 1.14 for "non-experts" on the log scale and 1.06 for "experts" and 1.07 for "non-experts" on the natural scale (see Table \ref{tab:scores-experts}).  
For two week ahead forecasts of deaths, "experts" performed worse than "non-experts" in terms of WIS on the log scale (WIS relative to the combined crowd ensemble: 1.29 vs. 1.06), but better on the natural scale (1.03 vs. 1.14). Both the "expert"- and the "non-expert"-ensemble had similar proportions of $R_t$ forecasts (mean of 32\% for "experts" and 32.2\% for "non-experts" across cases and deaths together).

@@

% \paragraph{Effort to recruit participants}. 
Just like \citet{bosseComparingHumanModelbased2022} and \citet{farrowHumanJudgmentApproach2017}, this study struggled to retain a large number of participants. Focused public outreach efforts such as creating a dedicated website, announcing an official tournament, providing a public leaderboard, sending weekly emails with details on past performance and weekly announcements on Twitter, did noticeably increase participation compared to the previous study in Germany and Poland. Nevertheless, retaining participants beyond the initial recruitment proved challenging, and most forecasters only submitted a single forecast. \citet{mcandrewChimericForecastingCombining2022} had a higher number of participants, suggesting that making use of existing forecasting platforms that have access to a large existing user base and greater resources may be helpful in recruiting a larger number of participants, though these platforms lack the flexibility and software tooling to run a novel study of this kind in real-time as things stand.

@@

%WIS scores
For forecasts of cases two weeks ahead, performance of the human judgement forecasts was better than the European Forecast Hub ensemble in terms of WIS on the log scale, and worse in terms of WIS on the natural scale. This was linked to a tendency of the Hub ensemble to make lower case predictions, which led to lower overprediction penalties on the natural scale, but noticeably higher underprediction penalties on the log scale. 
For forecasts of deaths two weeks ahead, direct human forecasts and the combined crowd ensemble performed better than the Hub ensemble on the log scale. On the natural scale, the combined crowd ensemble performed worse than the Hub ensemble, while the direct crowd forecasts still performed better. $R_t$ forecasts for deaths performed noticeably worse than all other approaches both on the log and on the natural scale.
